A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2019-02-13 doc/large_strain.tex
Reinstate comment on lack of p-refinement for GMG
Add note on GMG framework / level operator comparison

{\color{red}
To summarize, in this section we propose a method by which to construct level tangent operators for finite-strain elasticity to be employed within the GMG preconditioner.
The other parts of the GMG preconditioner such as level smoothers, matrix-free interlevel transfer operators and the coarse level iterative solver are provided by the \texttt{deal.II} library.
The efficiency of the preconditioner with the considered level operators will be assessed by numerical examples in the next section.
We will, however, refrain from studying the machine performance aspects of the multigrid preconditioner as a whole as the focus of this paper is the matrix-free implementation of tangent operators of finite-strain elasticity both at the fine mesh level as well as the coarser multigrid levels.
% Finally, we note that at the time of writing, a $p-$ version of the GMG was not available within the \texttt{deal.II} library and therefore not considered here.
}

@@

%{\color{red}
%Additionally, we compare solution times of the preconditioned iterative solvers.
%Our goal is to \changeJP{determine} whether the matrix-free approach with a GMG preconditioner can be faster than the matrix-based approach with an AMG preconditioner.
%}
\changeJP{%
Additionally, we compare solution times of the preconditioned iterative solvers.
Our goal is to \changeJP{determine} whether the matrix-free approach with a GMG preconditioner can be faster than the matrix-based approach with a well-established implementation of an AMG preconditioner.
}
%\changeJP{We note that although the well-established implementation of the AMG preconditioner \cite{Gee2006a} is the best available choice for the matrix-based implementation of the boundary value problem, the current implementation of the matrix-free GMG preconditioner is not yet optimized and can be improved further.}
Figures \ref{fig:benchmark_miehe_Emmy_sol2}, \ref{fig:benchmark_miehe_Emmy_sol3}, \ref{fig:benchmark_miehe_IWR_sol2} and \ref{fig:benchmark_miehe_IWR_sol3} confirm that the efficiency of the proposed GMG preconditioner translates into faster solution times for {\color{red}all} polynomial degrees.
{\color{red}Most importantly, the wall time per DoF for the matrix-free solver with GMG preconditioner stays almost constant both for 2D and 3D problems.
\changeJP{In comparison}, the wall time of the matrix-based iterative solver grows rapidly with increasing polynomial degree.
Clearly this is related both to the performance of the preconditioner from the linear algebra perspective as well as the matrix-vector products, studied in the previous section.
Based on these results, we conclude that the matrix-free approach with the adopted here geometric multigrid preconditioner can deliver a very competitive solution strategy for \changeJP{engineering problems
with compressible hyperelastic} finite strain material models.
Extension of this study from the node level to the cluster level is beyond the scope of this work,
however we hypothesize that the matrix-free and multigrid frameworks provided by \texttt{deal.II} library would provide a competitive solution for that case as well.}

>>>>

{\color{red}
To summarize, in this section we propose a method by which to construct level tangent operators for finite-strain elasticity to be employed within the GMG preconditioner.
The other parts of the GMG preconditioner such as level smoothers, matrix-free interlevel transfer operators and the coarse level iterative solver are provided by the \texttt{deal.II} library.
The efficiency of the preconditioner with the considered level operators will be assessed by numerical examples in the next section.
We will, however, refrain from studying the machine performance aspects of the multigrid preconditioner as a whole as the focus of this paper is the matrix-free implementation of tangent operators of finite-strain elasticity both at the fine mesh level as well as the coarser multigrid levels.
Finally, we note that at the time of writing, a $p-$ version of the GMG was not available within the \texttt{deal.II} library and therefore not considered here.
}

@@

%{\color{red}
%Additionally, we compare solution times of the preconditioned iterative solvers.
%Our goal is to \changeJP{determine} whether the matrix-free approach with a GMG preconditioner can be faster than the matrix-based approach with an AMG preconditioner.
%}
\changeJP{%
Additionally, we compare solution times of the preconditioned iterative solvers.
Our goal is to \changeJP{determine} whether the matrix-free approach with a GMG preconditioner can be faster than the matrix-based approach with a well-established implementation of an AMG preconditioner.
We note specifically that the comparison is restricted to the choice of level operators used with the pre-existing GMG framework, and not the framework itself. 
}
%\changeJP{We note that although the well-established implementation of the AMG preconditioner \cite{Gee2006a} is the best available choice for the matrix-based implementation of the boundary value problem, the current implementation of the matrix-free GMG preconditioner is not yet optimized and can be improved further.}
Figures \ref{fig:benchmark_miehe_Emmy_sol2}, \ref{fig:benchmark_miehe_Emmy_sol3}, \ref{fig:benchmark_miehe_IWR_sol2} and \ref{fig:benchmark_miehe_IWR_sol3} confirm that the efficiency of the proposed GMG preconditioner translates into faster solution times for {\color{red}all} polynomial degrees.
{\color{red}Most importantly, the wall time per DoF for the matrix-free solver with GMG preconditioner stays almost constant both for 2D and 3D problems.
\changeJP{In comparison}, the wall time of the matrix-based iterative solver grows rapidly with increasing polynomial degree.
Clearly this is related both to the performance of the preconditioner from the linear algebra perspective as well as the matrix-vector products, studied in the previous section.
Based on these results, we conclude that the matrix-free approach with the adopted here geometric multigrid preconditioner can deliver a very competitive solution strategy for \changeJP{engineering problems
with compressible hyperelastic} finite strain material models.
Extension of this study from the node level to the cluster level is beyond the scope of this work,
however we hypothesize that the matrix-free and multigrid frameworks provided by \texttt{deal.II} library would provide a competitive solution for that case as well.}
